 

### 2020-09-29 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

Forest canopies moderate relative humidity in below-canopy microclimates along with temperature. In three European forest types (Broadleaved, Non-Pine Conifer, Pine) canopy buffering was observed to decrease daily maximum $T_{air}$ and increase daily minimum RH in below-canopy microclimate [@vonarxSpatiotemporalEffectsForest2012]

@@

**Finally, carbon dioxide (CO~2~) concentrations tend to be higher in the understory.** 
-	Higher in understory, particularly at dusk (Koike et al. 2001).
-	Higher in the understory overnight; difference persists during the day but is very small [@yang_spatial_1999].
Differences in concentration are by far most pronounced near ground level. 
Given that differences are small during the day when photosynthesis is active, and that even nighttime differences are modest at the height of understory tree crowns, CO~2~ concentration is unlikely to have much effect on the energy balance and metabolism of leaves across the forest vertical gradient.

@@

**The strength of this buffering varies across forests, being influenced by both forest characteristics and the biophysical environment.**
First, buffering increases with canopy cover.
* greater cover --> lower max T and VPD, higher minT  [@davis_microclimatic_2019]
* greater cover -->  lower max T and light availability [@zellweger_seasonal_2019]
* [@thom_effects_2020; Tymen et al. 2017 ; Jucker et al. 2018 ; Hardwick et al. 2015]
*Presumably, buffering would also be affected by canopy roughness, which affects turbulent air flow and the canopy boundary layer.*
Taller trees don’t necessarily increase buffering [@zellweger_seasonal_2019]. 
Species that have higher shade casting ability increase T buffering (Zelllweger et al. 2019) [@zellweger_seasonal_2019]
The strength of buffering also varies with respect to geographic and climatic factors.
* Distance to coast, topographic position, elevation [@zellweger_seasonal_2019]
* [@davis_microclimatic_2019]
However vertical climate gradients within forests are known to be steeper than those driven by elevation and latitude [@nakamuraForestsTheirCanopies2017,  @martinCovarianceSunShade2020a]

@@

Leaf traits vary along the vertical canopy profile following the pattern of light availability across the gradient [@poorterMetaanalysisPlantResponses2019b]. Leaves in the upper canopy that are exposed to more light differ in their traits compared to shaded leaves exposed to less light. Increasing number of studies point towards the importance of incorporating vertical canopy leaf trait variation in modelling for leaf economic spectrum (LES) [Chen et al. 2020], leaf thermoregular and thermal sensitivty which will provide a more accurate representation of within-canopy gradient response to increasing temperatures.[@fauset_differences_2018, @michaletzEnergeticCarbonEconomic2016, @niinemetsWorldwideAnalysisWithincanopy2015, Chen et al. 2020, REFS)

@@

>>>>

Forest canopies moderate relative humidity in below-canopy microclimates along with temperature. In three European forest types (Broadleaved, Non-Pine Conifer, Pine) canopy buffering was observed to decrease daily maximum $T_{air}$ and increase daily minimum RH in below-canopy microclimate.With rising temperatures, this trend in maximum $T_{air} and RH would have implications on understory seedling growth, especially in high altitude temperate pine forests [@vonarxSpatiotemporalEffectsForest2012]

@@

**Finally, carbon dioxide (CO~2~) concentrations tend to be higher in the understory.** 
-	Higher in understory, particularly at dusk (Koike et al. 2001).
-	Higher in the understory overnight; difference persists during the day but is very small [@yang_spatial_1999].
Differences in concentration are by far most pronounced near ground level.[@niinemetsPhotosyntheticAcclimationSimultaneous2004]
Given that differences are small during the day when photosynthesis is active, and that even nighttime differences are modest at the height of understory tree crowns, CO~2~ concentration is unlikely to have much effect on the energy balance and metabolism of leaves across the forest vertical gradient.

@@

**The strength of this buffering varies across forests, being influenced by both forest characteristics and the biophysical environment.**
First, buffering increases with canopy cover.
* greater cover --> lower max T and VPD, higher minT  [@davis_microclimatic_2019]
* greater cover -->  lower max T and light availability [@zellweger_seasonal_2019]
* [@thom_effects_2020; Tymen et al. 2017 ; Jucker et al. 2018 ; Hardwick et al. 2015]
*Presumably, buffering would also be affected by canopy roughness, which affects turbulent air flow and the canopy boundary layer.*
Taller trees don’t necessarily increase buffering [@zellweger_seasonal_2019]. 
Species that have higher shade casting ability such as braodleaved trees increase T buffering (Zelllweger et al. 2019) [@zellweger_seasonal_2019,@vonarxSpatiotemporalEffectsForest2012]
The strength of buffering also varies with respect to geographic and climatic factors.
* Distance to coast, topographic position, elevation [@zellweger_seasonal_2019]
* [@davis_microclimatic_2019]
However vertical climate gradients within forests are known to be steeper than those driven by elevation and latitude [@nakamuraForestsTheirCanopies2017,  @martinCovarianceSunShade2020a]

@@

Leaf traits vary along the vertical canopy profile following the pattern of light availability across the gradient [@poorterMetaanalysisPlantResponses2019b]. Leaves in the upper canopy that are exposed to more light differ in their traits compared to shaded leaves exposed to less light [@@niinemetsWorldwideAnalysisWithincanopy2015] Increasing number of studies point towards the importance of incorporating vertical canopy leaf trait variation in modelling for leaf economic spectrum (LES) [Chen et al. 2020], leaf thermoregular and thermal sensitivty which will provide a more accurate representation of within-canopy gradient response to increasing temperatures.[@fauset_differences_2018, @michaletzEnergeticCarbonEconomic2016, @niinemetsWorldwideAnalysisWithincanopy2015, Chen et al. 2020, REFS)

@@

Photosynthesis increases with increasing temperature and saturates at quatum flux density of 400 to 700nm m-2 s-1, beyond this the leaf experiences heat stress, undergoes non-photochemical quenching, photoinhibition and leaf necrosis [@niinemetsWorldwideAnalysisWithincanopy2015]
